Model-based pretreatment for regression with correlated variables.

### 2014-04-07 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@500 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn. Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even without bias from modelisation. 
In this paper, vectors and matrices are in bold letters to distinguish them from scalars. We note the linear regression model:
\begin{equation}
		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}
	\end{equation}
	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ the noise of the regression, with $\boldsymbol{I_n}$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$. The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS}). %As shown in section \ref{sectionOLS}, 
	Estimation of the parameters requires the inversion of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
Variance of $\boldsymbol{\hat\beta}$ increases based on two aspects :
\begin{itemize}
	\item The dimension $p$ (number of covariates) of the model: the more covariates you have the greater variance you get.
	\item The correlations within the covariates: strongly correlated covariates give bad-conditioning and increase variance of the estimators .
\end{itemize}

@@

	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty.
		\begin{equation}
		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\boldsymbol{\beta} \parallel_1\leq \lambda
		\end{equation}	
	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive. It requires only the same order of magnitude of computational effort as \textsc{OLS} applied to the full set of covariates. And it  really selects covariates with coefficients set exactly to 0.
	 But \textsc{LASSO} also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates. Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). Some recent variants of the \textsc{LASSO} does exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
	 \\

>>>>

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn. Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even without bias from modelisation. 
In this paper, vectors and matrices are in bold letters to distinguish them from scalars. We note the linear regression model:
\begin{equation}
		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}
	\end{equation}
	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ the noise of the regression, with $\boldsymbol{I_n}$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$. The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, that can be estimated with Ordinary Least Squares (\textsc{OLS}). %As shown in section \ref{sectionOLS}, 
	Estimation of the parameters requires the inversion of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
Variance of $\boldsymbol{\hat\beta}$ increases based on two aspects :
\begin{itemize}
	\item The dimension $p$ (number of covariates) of the model: the more covariates you have the greater variance you get.
	\item The correlations within the covariates: strongly correlated covariates give bad-conditioning and increase variance of the estimators .
\end{itemize}

@@

	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty to obtain zeros in $\hat{\boldsymbol{\beta}}$:
		\begin{equation}
		 \boldsymbol{\hat{\beta}}=\operatorname{argmin} \left\lbrace \parallel \boldsymbol{Y}-\boldsymbol{X\beta}\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\boldsymbol{\beta} \parallel_1\leq \lambda
		\end{equation}	
	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive. It requires only the same order of magnitude of computational effort as \textsc{OLS} applied to the full set of covariates. And it  really selects covariates with coefficients set exactly to 0.
	 But \textsc{LASSO} also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates. 
%	 Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). 
	 Some recent variants of the \textsc{LASSO} does exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random \textsc{LASSO} \cite{wang2011random}.
	 \\
